---
title: "Do the people of SMI belive in god or not?"
author: Meenakshi
categories: Assessment 3
description: analysis and EDA
format: html
---

# Peasant Hunch: Do the people of SMI belive in god or not?

a survey of people from SMI, as what their opinion on god existing is ( Yes / No).

## Data Dictionary

A tibble with 47 observations and 3 variables: \## Data dictionary

65 observations and 4 variables.

## Qualitative Variables:

> -   **gender** Gender of the student, M, F or Other.
> -   **name** Name of the student.
> -   **believe_in_god** do they believe in god (yes,no).

## SetUp R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(naniar)
library(visStatistics)
library(janitor)
library(ggformula)
library(ggridges)
library(skimr)
library(resampledata)
library(checkdown)
library(tinytable)
library(GGally)
library(corrplot)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(marquee)
library(corrgram)
library(crosstable)
library(kableExtra)
library(tinytable)
library(DT)
library(paletteer)
```

## Reading Data

```{r}
god <- readr::read_csv("Data/believe_in_god.csv")
```

## 🔍 Examine Data

```{r}
dim(god)
base::names(god)
utils::str(god)
god <- god %>% clean_names(case = "snake")
dplyr::glimpse(god)
```

## Dropping NA

```{r}
god_modified <- god %>%
  naniar::replace_with_na_all(condition = ~ .x %in% common_na_numbers) %>%
  replace_with_na_all(condition = ~ .x %in% common_na_strings)

god_modified <- god_modified %>% drop_na()
glimpse(god_modified)
```

## Viewing Missing Data

```{r}
visdat::vis_miss(god_modified)
visdat::vis_dat(god_modified)
```

## Converting to a factor

```{r}
god_modified <- god_modified %>%
  mutate(
    gender = as.factor(gender),
    name = as.factor(name),
    believe_in_god = as.factor(believe_in_god)
  )

glimpse(god_modified)
```

## 👾 Data Cleaned

```{r}
god_modified %>%
  DT::datatable(
    style = "default",
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: black; font-size: 100%;", "Do people believe in god or not"
    ),
    options = list(pageLength = 10, autoWidth = TRUE)
  ) %>%
  DT::formatStyle(
    columns = names(god_modified),
    fontFamily = "Georgia",
    fontSize = "12px"
  )
```

## prop test

```{r}
god_tbl <- table(god_modified$gender, god_modified$believe_in_god)

prop.test(god_tbl)
```

## grouped barchart of belief grouped by gender

```{r}
god_modified %>%
  count(gender, believe_in_god) %>%
  ggplot(aes(x = gender, y = n, fill = believe_in_god)) +
  geom_col(position = "dodge") +
  labs(
    title = "Count of Belief in God by Gender",
    x = "Gender",
    y = "Count",
    fill = "Believe in God"
  ) +
  scale_fill_manual(values = c("#5BA19B", "#F29E4C")) +
  theme_minimal(base_size = 15)

```

## mosaic plot of belief and gender

```{r}
mosaic(~ gender + believe_in_god,
       data = god_modified,
       shade = TRUE,  
       legend = TRUE, 
       labeling_args = list(
         set_varnames = c(gender = "Gender", believe_in_god = "Belief in God")
       ),
       main = "Belief in God by Gender")

```
