---
title: "Weekly Expenditure of students in SMI"
author: Meenakshi
categories: Assessment 3
description: analysis and EDA
format: html
---

# Free Hunch #1: Guys have more pocket money than Gals.

Have you got money? Open your wallet and show me how much. Guys and Gals. Will this work in these UPI days? And well-dressed Gals don’t have pockets in their clothing, Uncle…

Al right, then ask “How much did you spend today?” Again, Guys and Gals.

And then a two-independent-sample test for means. And Permutation and all that stuff.

## Data Dictionary

A tibble with 40 observations and 3 variables:

## Qualitative Variables:

> -   **gender** Gender of the student, M, F or Other.
> -   **name** Name of the student. 

## Quantitative Variables:
> -   **Total_Expenditure_Last_Week** Total Expenditure Last Week (in numbers).

## SetUp R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(naniar)
library(visStatistics)
library(janitor)
library(ggformula)
library(ggridges)
library(skimr)
library(resampledata)
library(checkdown)
library(tinytable)
library(GGally)
library(corrplot)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(marquee)
library(corrgram)
library(crosstable)
library(kableExtra)
library(tinytable)
library(DT)
library(paletteer)
```

## Reading Data

```{r}
money <- readr::read_csv("Data/weekly_expenditure.csv")
```

## 🔍 Examine Data

```{r}
dim(money)
base::names(money)
dplyr::glimpse(money)
utils::str(money)
money <- money %>% clean_names(case = "snake")
```

## Dropping NA

```{r}
money_modified <- money %>%
  naniar::replace_with_na_all(condition = ~ .x %in% common_na_numbers) %>%
  replace_with_na_all(condition = ~ .x %in% common_na_strings)

money_modified <- money_modified %>% drop_na()
glimpse(money_modified)
```

## Viewing Missing Data

```{r}
visdat::vis_miss(money_modified)
visdat::vis_dat(money_modified)
```

## Converting to a factor

```{r}
money_modified <- money_modified %>%
  mutate(
    name = as.factor(name)
  )

glimpse(money_modified)
```

## 👾 Data Cleaned

```{r}
money_modified %>%
  DT::datatable(
    style = "default",
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: black; font-size: 100%;", "Weekly Expenditure"
    ),
    options = list(pageLength = 10, autoWidth = TRUE)
  ) %>%
  DT::formatStyle(
    columns = names(money_modified),
    fontFamily = "Georgia",
    fontSize = "12px"
  )
```

## Ttest

```{r}
library(dplyr)
money_modified$total_expenditure_last_week <- as.numeric(as.character(money_modified$total_expenditure_last_week))
# Two-sample t-test
t.test(total_expenditure_last_week ~ gender, data = money_modified)

# Permutation test setup
set.seed(123)
obs_diff <- money_modified %>%
  group_by(gender) %>%
  summarize(mean_spend = mean(total_expenditure_last_week)) %>%
  summarize(diff = diff(mean_spend)) %>%
  pull(diff)

perm_diff <- function(data, n_perm = 1000){
  replicate(n_perm, {
    permuted_gender <- sample(data$gender)
    data %>%
      mutate(gender = permuted_gender) %>%
      group_by(gender) %>%
      summarize(mean_spend = mean(total_expenditure_last_week)) %>%
      summarize(diff = diff(mean_spend)) %>%
      pull(diff)
  })
}

permutation_diffs <- perm_diff(money_modified)

# Calculate two-tailed permutation p-value
mean(abs(permutation_diffs) >= abs(obs_diff))
```

```{r}
ggplot(money_modified, aes(x = total_expenditure_last_week, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Density Plot of Weekly Expenditure by Gender",
    x = "Total Expenditure Last Week",
    y = "Density"
  ) +
  scale_fill_manual(values = c("#5BA19B", "#F29E4C")) +
  theme_minimal(base_size = 15)

```
