---
title: "Tattoo Attraction in SMI"
author: Meenakshi
categories: Assessment 3
description: analysis and EDA
format: html
---
# Free Hunch #10: Do tattoos make people look attractive or not?
 a survey of people from different colleges on campus, as what their opinion is on tattoos and attractiveness ( Yes / No). Survey Males and Females, Students and Faculty. Check the number of respondents in each sub-group!! And then a chi-square test to see if there are differences in proportions.


## Data Dictionary

A tibble with 40 observations and 3 variables:

## Qualitative Variables:

> -   **gender** Gender of the student, M, F or Other.
> -   **Tattoo_Attractive** Do they find tattoos attractive (yes/no)
> -   **name** Name of the student.
## SetUp R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(naniar)
library(visStatistics)
library(janitor)
library(ggformula)
library(ggridges)
library(skimr)
library(resampledata)
library(checkdown)
library(tinytable)
library(GGally)
library(corrplot)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(marquee)
library(corrgram)
library(crosstable)
library(kableExtra)
library(tinytable)
library(DT)
library(paletteer)
```

## Reading Data

```{r}
tattoo <- readr::read_csv("Data/tattoo_attraction.csv")
```

## 🔍 Examine Data

```{r}
dim(tattoo)
base::names(tattoo)
dplyr::glimpse(tattoo)
utils::str(tattoo)
tattoo <- tattoo %>% clean_names(case = "snake")
```

## Dropping NA

```{r}
tattoo_modified <- tattoo %>%
  naniar::replace_with_na_all(condition = ~ .x %in% common_na_numbers) %>%
  replace_with_na_all(condition = ~ .x %in% common_na_strings)

tattoo_modified <- tattoo_modified %>% drop_na()
glimpse(tatoo_modified)
```

## Viewing Missing Data

```{r}
visdat::vis_miss(tattoo_modified)
visdat::vis_dat(tattoo_modified)
```

## Converting to a factor

```{r}
tattoo_modified <- tattoo_modified %>%
  mutate(
    gender = as.factor(gender),
    name = as.factor(name),
    tattoo_attractive = as.factor(tattoo_attractive)
  )

glimpse(tattoo_modified)
```

## 👾 Data Cleaned

```{r}
tattoo_modified %>%
  DT::datatable(
    style = "default",
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: black; font-size: 100%;", "Do tattoos make people look attractive or not"
    ),
    options = list(pageLength = 10, autoWidth = TRUE)
  ) %>%
  DT::formatStyle(
    columns = names(tattoo_modified),
    fontFamily = "Georgia",
    fontSize = "12px"
  )
```
## Do students in SMI find tattoos attractive
```{r}
tattoo_modified %>%
  count(tattoo_attractive) %>%
  ggplot(aes(x = tattoo_attractive, y = n, fill = tattoo_attractive)) +
  geom_col(width = 0.7) +
  labs(
    title = "Are Tattoos Considered Attractive?",
    x = "",
    y = "Number of Responses"
  ) +
  scale_fill_manual(values = c("#5BA19B", "#F29E4C")) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "none")
```
## Tattoo atractiveness grouped by gender
```{r}
tattoo_modified %>%
  count(gender, tattoo_attractive) %>%
  ggplot(aes(x = gender, y = n, fill = tattoo_attractive)) +
  geom_col(position = "dodge", width = 0.7) +
  labs(
    title = "Are Tattoos Considered Attractive? (Grouped by Gender)",
    x = "Gender",
    y = "Number of Responses"
  ) +
  scale_fill_manual(values = c("#5BA19B", "#F29E4C")) +
  theme_minimal(base_size = 15)
```

## chisquare test
```{r}
# Creating a contingency table
tattoo_gender_tab <- table(tatoo_modified$gender, tatoo_modified$tattoo_attractive)

# Runing Chi-square test
chisq.test(tattoo_gender_tab)

```

