---
title: "Friends vs Modern Family vs House vs Big Bang Theory"
author: Meenakshi
categories: Assessment 3
description: analysis and EDA
format: html
---

# Free Hunch #4b:  Friends vs Modern Family vs House vs Big Bang Theory

And then a two-independent-sample test for means. And Permutation and all that stuff.

Sure! Based on your **`tv_series.csv`** file, here’s the data dictionary written in the same format and tone as your example 👇

---

## Data Dictionary

A tibble with 52 observations and 5 variables:

## Qualitative Variables:

>  **name**
>   Name of the student respondent.
>
>  **gender**
>   Gender of the student — Male or Female.

## Quantitative Variables:

>  **friends**
>   Rating given by the student for the TV show *Friends* (on a numerical scale).
>
>  **modern_family**
>   Rating given by the student for the TV show *Modern Family* (on a numerical scale).
>
>  **big_bang_theory**
>   Rating given by the student for the TV show *The Big Bang Theory* (on a numerical scale).
 
## SetUp R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(naniar)
library(visStatistics)
library(janitor)
library(ggformula)
library(ggridges)
library(skimr)
library(resampledata)
library(checkdown)
library(tinytable)
library(GGally)
library(corrplot)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(marquee)
library(corrgram)
library(crosstable)
library(kableExtra)
library(tinytable)
library(DT)
library(paletteer)
```

## Reading Data

```{r}
series <- readr::read_csv("Data/tv_series.csv")
```

## 🔍 Examine Data

```{r}
dim(series)
base::names(series)
dplyr::glimpse(series)
utils::str(series)
series <- series %>% clean_names(case = "snake")
```

## Dropping NA

```{r}
series_modified <- series %>%
  naniar::replace_with_na_all(condition = ~ .x %in% common_na_numbers) %>%
  replace_with_na_all(condition = ~ .x %in% common_na_strings)

series_modified <- series_modified %>% drop_na()
glimpse(series_modified)
```

## Viewing Missing Data

```{r}
visdat::vis_miss(series_modified)
visdat::vis_dat(series_modified)
```

## Converting to a factor

```{r}
series_modified <- series_modified %>%
  mutate(
    name = as.factor(name),
    gender = as.factor(gender),
  )

glimpse(money_modified)
```

## 👾 Data Cleaned

```{r}
series_modified %>%
  DT::datatable(
    style = "default",
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: black; font-size: 100%;", "Friends vs Modern Family vs House vs Big Bang Theory"
    ),
    options = list(pageLength = 10, autoWidth = TRUE)
  ) %>%
  DT::formatStyle(
    columns = names(series_modified),
    fontFamily = "Georgia",
    fontSize = "12px"
  )
```
## ANOVA
```{r}
## ANOVA for Friends
anova_friends <- aov(friends ~ gender, data = series_modified)
summary(anova_friends)

## ANOVA for Modern Family
anova_modern <- aov(modern_family ~ gender, data = series_modified)
summary(anova_modern)

## ANOVA for Big Bang Theory
anova_bigbang <- aov(big_bang_theory ~ gender, data = series_modified)
summary(anova_bigbang)
```

##Boxplot
```{r}
ggplot(series_modified, aes(x = gender, y = friends, fill = gender)) +
  geom_boxplot() +
  labs(title = "Friends Ratings by Gender", x = "Gender", y = "Rating") +
  theme_minimal()
```

