---
title: "Weird Dad"
author: Meenakshi
categories: Assessment 3
description: analysis and EDA
format: html
---

# Free Hunch #8: All Srishti Dads are Weird.

Here's a random sampling of finding whether Srishti dad's are weird or not, we also venture onto MIT dads, showing two types demographics here.

## Data dictionary

65 observations and 4 variables.

## Qualitative Variables:

> -   **gender** Gender of the student, M, F or Other.
> -   **college** College of the student, either being SMI or MIT.
> -   **is_dad_weird** Student perception of whether they found their dad weird or not.
> -   **name** Name of the student.

## SetUp R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(naniar)
library(visStatistics)
library(janitor)
library(ggformula)
library(ggridges)
library(skimr)
library(resampledata)
library(checkdown)
library(tinytable)
library(GGally)
library(corrplot)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(marquee)
library(corrgram)
library(crosstable)
library(kableExtra)
library(tinytable)
library(DT)
library(paletteer)
```

## Reading Data

```{r}
dad_weird <- readr::read_csv("Data/weird_dads.csv")
```

## 🔍 Examine Data

```{r}
dim(dad_weird)
base::names(dad_weird)
dplyr::glimpse(dad_weird)
utils::str(dad_weird)
```

## Dropping NA

```{r}
dad_weird_modified <- dad_weird %>%
  naniar::replace_with_na_all(condition = ~ .x %in% common_na_numbers) %>%
  replace_with_na_all(condition = ~ .x %in% common_na_strings)

dad_weird_modified <- dad_weird_modified %>% drop_na()
glimpse(dad_weird_modified)
```

## Viewing Missing Data

```{r}
visdat::vis_miss(dad_weird_modified)
visdat::vis_dat(dad_weird_modified)
```

## Converting to a factor

```{r}
dad_weird_modified <- dad_weird_modified %>%
  mutate(
    gender = as.factor(gender),
    college = as.factor(college),
    is_dad_weird = as.factor(is_dad_weird),
    name = as.factor(name)
  )

glimpse(dad_weird_modified)
```

## 👾 Data Cleaned

```{r}
dad_weird_modified %>%
  DT::datatable(
    style = "default",
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: black; font-size: 100%;", "Weird Dads in SMI"
    ),
    options = list(pageLength = 10, autoWidth = TRUE)
  ) %>%
  DT::formatStyle(
    columns = names(dad_weird_modified),
    fontFamily = "Georgia",
    fontSize = "12px"
  )
```

## Mosaic plots 001

```{r}
# MIT
mit_data <- dad_weird_modified %>% filter(college == "MIT")
mosaicplot(~ gender + is_dad_weird, data = mit_data, color = TRUE, main = "MIT: Gender to Weird Dad")

# SMI
smi_data <- dad_weird_modified %>% filter(college == "SMI")
mosaicplot(~ gender + is_dad_weird, data = smi_data, color = TRUE, main = "SMI: Gender vs Weird Dad")

```

## Statistical Analysis

```{r}
tab <- table(dad_weird_modified$college, dad_weird_modified$is_dad_weird)
print(tab)
```

## Chi-square test

```{r}
chisq.test(tab)
```

## Mosaic plots 002

```{r}


tab <- table(dad_weird_modified$college, dad_weird_modified$is_dad_weird)


mosaic(~ college + is_dad_weird, data = dad_weird_modified,
       shade = TRUE,
       main = "Are dads weird?: MIT vs SMI")
```

## Bar chart

```{r}
gf_bar(~is_dad_weird | college, data = dad_weird_modified, fill = ~is_dad_weird, position = "fill")

```
