---
title: "A2 Cardio-Dataset"
author: "Swetha KV"
date: "2025-10-12"
categories: [Scottish CVD]
format: html
---

## Setting up R Packages

```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(naniar)
library(skimr)
library(crosstable)
library(tinytable)
library(ggformula)
library(mosaic)
library(tidyverse)
library(vcd)
library(vcdExtra) 
library(resampledata)
library(ggmosaic)
library(ggpubr) 
library(visdat)
library(naniar) 
library(DT)
library(ggrepel)
library(marquee)
library(visStatistics)
```

## Read Data

```{r}
data(package = "CardioDataSets")
```

```{r}
data("scottish_CVD_df", package = "CardioDataSets")
View( scottish_CVD_df )
```

```{r}
data("scottish_CVD_df", package = "CardioDataSets")
dplyr::glimpse( scottish_CVD_df )
```

##Data Dictionary


##Examine Data

```{r}
base::names(scottish_CVD_df)
```

```{r}
dplyr::glimpse(scottish_CVD_df)
```

```{r}
base::dim(scottish_CVD_df)
```

```{r}
utils::str(scottish_CVD_df)
```

## Viewing Missing Data

```{r}
visdat::vis_dat(scottish_CVD_df)
```

```{r}
visdat::vis_miss(scottish_CVD_df)
```

```{r}
scottish_CVD_df <- scottish_CVD_df %>% 
 tidyr::drop_na()
```

```{r}
scottish_CVD_df <- scottish_CVD_df %>%
  naniar::replace_with_na_all(data = ., condition = ~ .x %in% common_na_numbers) %>%
  naniar::replace_with_na_all(data = ., condition = ~ .x %in% common_na_strings)
glimpse(scottish_CVD_df_modified)
```

```{r}
visdat::vis_miss(scottish_CVD_df)
```

```{r}
scottish_CVD_df %>% tt ()
```
```{r}
scottish_CVD_df %>%
  DT::datatable(
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: black; font-size: 150%;",
      "GSS Wages Dataset (Clean)"
    ),
    options = list(pageLength = 10, autoWidth = TRUE)
  ) %>%
  DT::formatStyle(
    columns = names(scottish_CVD_df),
    fontFamily = "Roboto Condensed",
    fontSize = "12px"
  )
```

```{r}
scottish_CVD_df <- scottish_CVD_df %>%
  mutate(across(where(is.character), as.factor)) %>% 
   mutate(cvddef = as.factor(cvddef)) %>% 
  relocate(where(is.factor))
glimpse(scottish_CVD_df)
```

```{r}
###Is there a relationship between age and socioeconomic deprivation (Carstairs index)?
mosaic::cor_test(age ~ carstair, data = scottish_CVD_df) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 3,
    caption = "Correlation between Age and Carstairs Index"
  )

```
There is no significant correlation between age and socioeconomic deprivation (Carstairs index) in the dataset (r = -0.003, p = 0.761).

This suggests that people of different ages are equally distributed across various levels of deprivation — older people are not more likely to live in deprived or affluent areas.

```{r}
scottish_CVD_df %>%
  drop_na(age, carstair) %>%
  gf_point(carstair ~ age) %>%
  gf_smooth() %>%
  gf_labs(
    title = "Age vs Carstairs Deprivation",
    subtitle = "Are older individuals from more deprived areas?",
    x = "Age (years)",
    y = "Carstairs Deprivation Index"
  )
```

```{r}
mosaic::cor_test(as.numeric(cvddef) ~ age, data = scottish_CVD_df) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 3)
```
The correlation (r = 0.362) is positive and moderate in strength.

The p-value = 0 means this result is highly significant - it’s not due to random chance.

The confidence interval (0.34 to 0.38) is narrow and fully positive → this strengthens our confidence.

As age increases, the chance of having cardiovascular disease (CVD) also increases.

```{r}
#1
scottish_CVD_df %>%
  drop_na(age, carstair, cvddef) %>%
  gf_point(carstair ~ age | cvddef, colour = ~cvddef) %>%
  gf_smooth(method = "lm", colour = "black") %>%
  gf_labs(
    title = "Age vs Deprivation by CVD Status",
    subtitle = "Older and more deprived individuals show higher CVD presence",
    x = "Age (years)",
    y = "Carstairs Index"
  )
```
```{r}
mosaic::cor_test(as.numeric(cvddef) ~ carstair, data = scottish_CVD_df)
```
```{r}
mosaic::cor_test(as.numeric(cvddef) ~ carstair, data = scottish_CVD_df) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 3,
    caption = "Correlation between CVD and Carstairs Index"
  )
##remove
```

```{r}
scottish_CVD_df <- scottish_CVD_df %>%
  mutate(smoke_num = as.numeric(smoke))  # nevr=1, mod=2, ex=3, hvy=4

mosaic::cor_test(as.numeric(cvddef) ~ smoke_num, data = scottish_CVD_df) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 3,
    caption = "Correlation between Smoking and CVD"
  )
##remove
```

```{r}
scottish_CVD_df <- scottish_CVD_df %>%
  mutate(sex_num = as.numeric(sex))  # female=1, male=2

mosaic::cor_test(age ~ sex_num, data = scottish_CVD_df) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 3,
    caption = "Correlation between Age and Sex"
  )
##remove
```

```{r}
scottish_CVD_df <- scottish_CVD_df %>%
  mutate(smoke_num = recode(smoke,
                            "nevr" = 0,
                            "ex"   = 1,
                            "mod"  = 2,
                            "hvy"  = 3))
mosaic::cor_test(smoke_num ~ age, data = scottish_CVD_df) %>%  
  broom::tidy() %>%
  knitr::kable(
    digits = 3,
    caption = "Correlation between Age and Sex")
##remove
```
```{r}
#2
scottish_CVD_df %>% 
  gf_boxplot(age ~ smoke, orientation = "x", fill = ~smoke, color = "black") %>%
  gf_labs(
    y = "Smoking Status",
    x = "Age of Individuals", 
    title = "Distribution of Age by Smoking Status",
    subtitle = "Checking if smoking habits vary by age"
  )

```
```{r}
#3 

mosaic::cor_test(carstair ~ smoke_num, data = scottish_CVD_df) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 3,
    caption = "Correlation between Deprivation and Smoking"
  )
```

