---
title: "Classwork 10"
author: Meenakshi
categories: Classwork
description: Learning to test for a single proportion, and inference test for two proportions.
format: html
---

## SetUp R Packages

```{r}
library(tidyverse)
library(mosaic) # Our go-to package
library(ggformula)
library(infer) # An alternative package for inference using tidy data
library(broom) # Clean test results in tibble form
library(skimr) # data inspection
library(tinytable) # Pretty Tables
library(kableExtra) # Pretty Tables
library(ggprism) # Prism-like ggplot themes
##
library(visStatistics) # All-in-one Stats test package
##
library(openintro) # datasets

library(vcd)
### Dataset from Chihara and Hesterberg's book (Second Edition)
library(resampledata)
```

## Case Study 1: YRBSS Survey

```{r}
data(yrbss, package = "openintro")
yrbss
```

For our first Research Question, we will choose the column `helmet_12m`: Remember that you can use `filter` to limit the dataset to just non-helmet wearers. Here, we will name the (filtered ) dataset `no_helmet`.

```{r}
yrbss %>%
  group_by(helmet_12m) %>%
  count()
```

Also, it may be easier to calculate the proportion if we create a new variable that specifies whether the individual has [*texted every day while driving*]{.underline} over the past 30 days or not. We will call this variable `text_ind`.

```{r}
yrbss %>%
  group_by(text_while_driving_30d) %>%
  count()
```

This is the proportion of people in this sample who do text when they drive without a helmet.

```{r}
no_helmet_text <- yrbss %>%
  filter(helmet_12m == "never") %>%
  mutate(text_ind = ifelse(text_while_driving_30d == "30", "yes", "no")) %>%
  # removing most of the other variables
  select(age, gender, text_ind)
no_helmet_text
```

Dropping NA and filtering to take count of the variable `text_ind` only.

```{r}
no_helmet_text %>%
  drop_na() %>%
  count(text_ind)
```

```{r}
no_helmet_text %>%
  drop_na() %>%
  summarize(prop = prop(text_ind, success = "yes"), n = n())
```

### Visualizing a Single Proportion

```{r}
ggplot2::theme_set(new = theme_classic())

no_helmet_text %>%
  drop_na() %>%
  gf_bar(~text_ind) %>%
  gf_labs(
    x = "texted?",
    title = "High-Schoolers who texted every day",
    subtitle = "While driving with no helmet on!!"
  )
```

### Inference

```{r}
mosaic::binom.test(~text_ind, data = no_helmet_text, success = "yes")
```

```{r}
mosaic::binom.test(~text_ind, data = no_helmet_text, success = "yes") %>%
  broom::tidy()
```

## Case Study 2: `GSS2002` dataset

```{r}
data(GSS2002, package = "resampledata")
glimpse(GSS2002)
```

We pull out Data of the variable `Education` and `DeathPenalty` only.

```{r}
GSS2002 %>% count(Education)
```

```{r}
GSS2002 %>% count(DeathPenalty)
```

### Data Munging

We drop NA entries in `Education` and `Death Penalty` and also check the levels of the two factors and set them up as we need:

```{r}
GSS2002_modified <- GSS2002 %>%
  dplyr::select(Education, DeathPenalty) %>%
  tidyr::drop_na(., c(Education, DeathPenalty)) %>%
  # Re-level the factors
  mutate(
    Education = factor(Education,
      levels = c("Left HS", "HS", "Jr Col", "Bachelors", "Graduate"),
      labels = c("Left HS", "HS", "Jr Col", "Bachelors", "Graduate"),
      # ordered = TRUE # Nope can't use this. See below
    ),
    DeathPenalty = factor(DeathPenalty,
      levels = c("Favor", "Oppose"),
      labels = c("Favor", "Oppose")
      # ordered = TRUE # Nope can't use this. See below
    )
  )
glimpse(GSS2002_modified)
```
### Contingency Table
```{r}
gss_vcd_table <- vcd::structable(Education ~ DeathPenalty, # cols ~ rows
  data = GSS2002_modified
)
gss_vcd_table %>% addmargins()
```
### Contingency Table Plots
```{r}
# Already computed
# gss_vcd_table <- vcd::structable(Education ~ DeathPenalty, # cols ~ rows
#                              data = GSS2002_modified)

gss_vcd_table %>%
  vcd::mosaic(
    gp = shading_hsv, direction = "h",
    main = "GSS2002 Education vs Death Penalty Mosaic Chart",
    legend = TRUE,
    labeling = labeling_border(
      varnames = c("F", "F"), # Remove variable name labels
      rot_labels = c(90, 0, 0, 0), # t,r,b,l?
      just_labels = c(
        "left", # Top Side. How?
        "left", # Right side
        "left", # Bottom side
        "right"
      )
    )
  ) # Left side. How?
```




























